---
title: "Stats 101C - Regression Project"
author: "Instructions"
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies: ["amsmath"]
  html_document: default
---
```{r, echo = FALSE, message=FALSE}
library(knitr)
library(tidyverse)
library(tidymodels)
library(stringr)
```


# Gameplan:

  - vfold CV the training data
  - Try multiple models: linear, multiple linear, tree, random forest, boosted tree, 
  - try multiple recipes: transformations on data
  - make models tunable with different parameters
  - make workflows with the different combinations of recipes and models AND different parameters
  - top 5 best result on assessment data will move on to the testing data
  - find best combination of model, recipe, parameters for the testing data


Getting initial glimpse of the data:

```{r}

# input into terminal the filepath to the Git project folder (containing train.csv)
x <- readline()
filepath <- gsub("\\\\","/",x)
trainFilepath <- paste0(filepath,"/train.csv")
testFilepath <- paste0(filepath,"/test.csv")
columnDescFilepath <- paste0(filepath,"/column_descriptions.csv")

train <- read.csv(trainFilepath)
test <- read.csv(testFilepath)
column_desc <- read.csv(columnDescFilepath)


```

```{r personal insights, message=F, eval=F}

glimpse(train)
glimpse(test)
view(column_desc)

```

```{r visualization of data}



```