---
title: "Stats 101C - Regression Project"
author: "Instructions"
output:
  pdf_document:
    latex_engine: xelatex
    extra_dependencies: ["amsmath"]
  html_document: default
---
```{r, echo = FALSE, message=FALSE}
# INSTALL ALL THESE PACKAGES
library(knitr)
library(tidyverse)
library(tidymodels)
library(stringr)
library(corrplot)
library(yardstick)
library(ranger)
library(xgboost)
```


# Gameplan:

  - vfold CV the training data
  - Try multiple models: multiple linear, decision tree, random forest, boosted tree, 
  - try multiple recipes: transformations on data
  - make models tunable with different parameters
  - make workflows with the different combinations of recipes and models with some parameters
  - find best model
  - fine tune the best model with more parameters
  - top 5 best result on assessment data will move on to the testing data
  - find lowest RMSE and highest RSQ 


Getting initial glimpse of the data:

```{r}

# SET THE WD TO YOUR FOLDER

trainFilepath <- paste0(getwd(),"/train.csv")
testFilepath <- paste0(getwd(),"/test.csv")
columnDescFilepath <- paste0(getwd(),"/column_descriptions.csv")

train <- read.csv(trainFilepath)
test <- read.csv(testFilepath)
column_desc <- read.csv(columnDescFilepath)

```

```{r personal insights, message=F, eval=F}

glimpse(train)
glimpse(test)
view(column_desc)

summary(train)

ncol(train)
nrow(train)

```

Response variable has a skewed right distribution (median is on the lower end). Possibly should normalize the response variable.

Data has big min's and max's, so there might be some need for sample stratification OR transformation. Most likely a log transformation because the predictor is percentage, so the range is only from 0-1.

```{r transforming data via recipes}

# remove name from training data
train$name <- NULL

# possible transformations: log, normalization, boxcox

# offset by 1 because there are some 0's in the data

# remove na observations

# remove all highly correlated data to avoid multicollinearity that would affect ML models

# NOTE: ALL OF THESE RECIPES OFFSET THE DATA FIRST BY 1 BEFORE TRANSFORMING

# TUNABLE PARAMS: threshold
# rec <- recipe(percent_dem ~ ., data = train %>% select(-name)) %>% 
#   step_rm(id) %>% 
#   step_mutate(pct_male = x0002e / x0001e) %>%
#   step_mutate(pct_female = x0003e / x0001e) %>%
#   step_mutate()
#   step_corr(all_numeric_predictors(), threshold = tune()) %>% 
#   step_mutate_at(all_numeric_predictors(), fn = function(x){x + 1}) %>% 
#   step_impute_knn(all_numeric_predictors())

# NOTE: we should probably use other methods of imputation

# rec_norm <- rec_knn %>% step_normalize(all_outcomes())

```

```{r v fold}

# CROSS VALIDATE TRAIN TO AVOID OVERFITTING

# set seed for reproducibility
set.seed(101)

# train_split <- vfold_cv(train, v = 5)

# make strata be response variable
train_split <- vfold_cv(train, v = 5, strata = percent_dem)

```

```{r model creation}

linear_model <- linear_reg() %>% set_engine("lm") %>% set_mode("regression")

# TUNABLE PARAMS: penalty, mixture -- for generalized linear model
regularized_model <- linear_reg(penalty = tune(), mixture = tune()) %>%
  set_engine("glmnet")
              
# TUNABLE PARAMS: cost_complexity, tree depth
tree_model <- decision_tree(cost_complexity = tune()) %>% 
  set_engine("rpart") %>% 
  set_mode("regression")

# TUNABLE PARAMS: mtry, min n for each node
r_forest_model <- rand_forest(trees = tune(), mtry = tune(), min_n = tune()) %>% 
  set_engine("ranger") %>% 
  set_mode("regression")
# random forest creates missing values for some reason when na's are ommitted, so imputation is used instead

# TUNABLE PARAMS: tree num, min n for each node, learn rate, loss reduction, tree depth
boost_tree_model <- boost_tree(learn_rate = tune(), loss_reduction = tune(), tree_depth = tune()) %>% 
  set_engine("xgboost") %>% 
  set_mode("regression")

# TUNABLE PARAMS: cost, margin -- support vector machines
svm_model <- svm_poly(cost = tune(), margin = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("regression")

# TUNABLE PARAMS: num_terms, prod_degree -- multivariate adaptive regression splines
mars_model <- mars(num_terms = tune(), prod_degree = tune()) %>%
  set_engine("earth") %>%
  set_mode("regression")

# TUNABLE PARAMS: hidden_units, penalty -- neural network ensamble 
## Note: engine would not be recognized
# nn_stack <- bag_mlp(hidden_units = tune(), penalty = tune()) %>%
#   set_mode("regression")
#   

model_list <- list(linear_model, regularized_model, tree_model, r_forest_model, boost_tree_model, svm_model, mars_model)



```

Possible models: linear, decision tree, random forest, boosted tree

```{r workflow creation for default recipe}

# make list of recipes and models to combine different permutations in workflows
# make them work for CV resampling

### Eric: Using three recipes: simple_rec, filter_rec, norm_rec -- all based on percentage population. Attempted to use spline recipes on significant variables based on linear regression, but the variables couldn't be recognized.


### Made simple_rec
simple_rec <- recipe(percent_dem ~ ., data = train) %>% 
  step_mutate(pct_male = x0002e / x0001e * 100) %>%
  step_mutate(pct_female = x0003e / x0001e * 100) %>%
  step_mutate(pct_under15 = (x0005e + x0006e + x0007e) * 100 / x0001e) %>%
  step_mutate(pct_15thru19 = x0008e * 100 / x0001e) %>%
  step_mutate(pct_20thru24 = x0009e * 100 / x0001e) %>%
  step_mutate(pct_25thru34 = x0010e * 100 / x0001e) %>%
  step_mutate(pct_35thru44 = x0011e * 100 / x0001e) %>%
  step_mutate(pct_45thru54 = x0012e * 100 / x0001e) %>%
  step_mutate(pct_55thru64 = (x0013e + x0014e) * 100 / x0001e) %>%
  step_mutate(pct_65thru84 = (x0015e + x0016e) * 100 / x0001e) %>%
  step_mutate(pct_over85 = (x0017e) * 100 / x0001e) %>%
  step_rename(median_age = x0018e) %>%
  step_mutate(pct_u18 = x0019e * 100 / x0001e) %>%
  step_mutate(pct_o16 = x0020e * 100 / x0001e) %>%
  step_mutate(pct_o18 = x0021e * 100 / x0001e) %>%
  step_mutate(pct_o21 = x0022e * 100 / x0001e) %>%
  step_mutate(pct_o62 = x0023e * 100 / x0001e) %>%
  step_mutate(pct_o65 = x0024e * 100 / x0001e) %>%
  step_mutate(pct_m_over18 = x0026e * 100 / x0001e) %>%
  step_mutate(pct_f_over18 = x0027e * 100 / x0001e) %>%
  step_mutate(pct_m_over65 = x0030e * 100 / x0001e) %>%
  step_mutate(pct_f_over65 = x0031e * 100 / x0001e) %>%
  step_mutate(pct_1race = x0034e * 100 / x0001e) %>%
  step_mutate(pct_over1race = x0035e * 100 / x0001e) %>%
  step_mutate(pct_white = x0037e * 100 / x0001e) %>%
  step_mutate(pct_black = x0038e * 100 / x0001e) %>%
  step_mutate(pct_natamer = (x0039e + x0040e + x0041e + x0042e + x0043e) * 100 / x0001e) %>%
  step_mutate(pct_asian = x0044e * 100 / x0001e) %>%
  step_mutate(pct_indian = x0045e * 100 / x0001e) %>%
  step_mutate(pct_chinese = x0046e * 100 / x0001e) %>%
  step_mutate(pct_filipino = x0047e * 100 / x0001e) %>%
  step_mutate(pct_japan = x0048e * 100 / x0001e) %>%
  step_mutate(pct_korean = x0049e * 100 / x0001e) %>%
  step_mutate(pct_viet = x0050e * 100 / x0001e) %>%
  step_mutate(pct_asian_other = x0051e * 100 / x0001e) %>%
  step_mutate(pct_pac_isl = x0052e * 100 / x0001e) %>%
  step_mutate(pct_nat_haw = x0053e * 100 / x0001e) %>%
  step_mutate(pct_chamorro = x0054e * 100 / x0001e) %>%
  step_mutate(pct_samoan = x0055e * 100 / x0001e) %>%
  step_mutate(pct_pacisl_other = x0056e * 100 / x0001e) %>%
  # step_mutate(pct_native_haw = (x0052e + x0053e + x0054e + x0055e + x0056e) * 100 / x0001e) %>%
  step_mutate(pct_otherrace = x0057e * 100 / x0001e) %>%
  step_mutate(pct_white_black = x0059e * 100 / x0001e) %>%
  step_mutate(pct_white_native = x0060e * 100 / x0001e) %>%
  step_mutate(pct_white_asian = x0061e * 100 / x0001e) %>%
  step_mutate(pct_black_native = x0062e * 100 / x0001e) %>%
  step_mutate(pct_whitecombo = x0064e * 100 / x0001e) %>%
  step_mutate(pct_blackcombo = x0065e * 100 / x0001e) %>%
  step_mutate(pct_nativecombo = x0066e * 100 / x0001e) %>%
  step_mutate(pct_asiancombo = x0067e * 100 / x0001e) %>%
  step_mutate(pct_hawcombo = x0068e * 100 / x0001e) %>%
  step_mutate(pct_othercombo = x0069e * 100 / x0001e) %>%
  step_mutate(pct_hispanic = x0071e * 100 / x0001e) %>%
  step_mutate(pct_mexican = x0072e * 100 / x0001e) %>%
  step_mutate(pct_puerto = x0073e * 100 / x0001e) %>%
  step_mutate(pct_cuban = x0074e * 100 / x0001e) %>%
  step_mutate(pct_hisp_other = x0075e * 100 / x0001e) %>%
  step_mutate(pct_not_hisp = x0076e * 100 / x0001e) %>%
  step_mutate(pct_nhisp_owhite = x0077e * 100 / x0001e) %>%
  step_mutate(pct_nhisp_oblack = x0078e * 100 / x0001e) %>%
  step_mutate(pct_nhisp_onatamer = x0079e * 100 / x0001e) %>%
  step_mutate(pct_nhisp_oasian = x0080e * 100 / x0001e) %>%
  step_mutate(pct_nhisp_ohawi = x0081e * 100 / x0001e) %>%
  step_mutate(pct_nhisp_other= x0082e * 100 / x0001e) %>%
  step_mutate(pct_nhisp_twoother = x0083e * 100 / x0001e) %>%
  step_rename(housing_units = x0086e) %>%
  step_mutate(pct_18t24_nohsdegree = c01_002e * 100 / c01_001e) %>%
  step_mutate(pct_18t24_hsdegree = c01_003e * 100 / c01_001e) %>%
  step_mutate(pct_18t24_somecollege = c01_004e * 100 / c01_001e) %>%
  step_mutate(pct_18t24_bachdegree = c01_005e * 100 / c01_001e) %>%
  step_mutate(pct_o25_nohsdegree = (c01_007e + c01_008e) * 100 / c01_006e) %>%
  step_mutate(pct_o25_hsdegree = c01_009e * 100 / c01_006e) %>%
  step_mutate(pct_o25_somecollege = (c01_010e + c01_011e) * 100 / c01_006e) %>%
  step_mutate(pct_o25_bachdegree = c01_012e * 100 / c01_006e) %>%
  step_mutate(pct_o25_graddegree = c01_013e * 100 / c01_006e) %>%
  step_mutate(pct_25t34_hsormore = c01_017e * 100 / c01_016e) %>%
  step_mutate(pct_25t34_bachormore = c01_018e * 100 / c01_016e) %>%
  step_mutate(pct_35t44_hsormore = c01_020e * 100 / c01_019e) %>%
  step_mutate(pct_35t44_bachormore = c01_021e * 100 / c01_019e) %>%
  step_mutate(pct_45t64_hsormore = c01_023e * 100 / c01_022e) %>%
  step_mutate(pct_45t64_bachormore = c01_024e * 100 / c01_022e) %>%
  step_mutate(pct_o65_hsormore = c01_026e * 100 / c01_025e) %>%
  step_mutate(pct_o65_bachormore = c01_027e * 100 / c01_025e) %>%
  step_impute_knn(income_per_cap_2016, income_per_cap_2017, income_per_cap_2018, income_per_cap_2019, income_per_cap_2020, gdp_2016, gdp_2017, gdp_2018, gdp_2019, gdp_2020, pct_18t24_nohsdegree, pct_18t24_hsdegree, pct_18t24_somecollege, pct_18t24_bachdegree) %>%
  step_rm(id, contains("x00"), contains("c01")) %>% 
  step_mutate_at(total_votes, median_age, housing_units, income_per_cap_2016, income_per_cap_2017, income_per_cap_2018, income_per_cap_2019, income_per_cap_2020, gdp_2016, gdp_2017, gdp_2018, gdp_2019, gdp_2020, fn = function(x){x + 1}) %>%
  step_log(total_votes, median_age, housing_units, income_per_cap_2016, income_per_cap_2017, income_per_cap_2018, income_per_cap_2019, income_per_cap_2020, gdp_2016, gdp_2017, gdp_2018, gdp_2019, gdp_2020, base = 10)
  
  # step_normalize(income_per_cap_2016, income_per_cap_2017, income_per_cap_2018, income_per_cap_2019, income_per_cap_2020, gdp_2016, gdp_2017, gdp_2018, gdp_2019, gdp_2020)

filter_rec <- simple_rec %>% 
  step_corr(all_numeric_predictors(), threshold = tune())

norm_rec <- simple_rec %>%
  step_normalize(all_numeric_predictors())


### The splines don't work -- not sure why -- they dont detect the variables in the cross-validation 


# simple_spline <- simple_rec %>%
#   step_ns(income_per_cap_2019, deg_free = tune("inc_2019 df")) %>%
#   step_ns(income_per_cap_2020, deg_free = tune("inc_2020 df")) %>%
#   step_ns(pct_o25_bachdegree, deg_free = tune("o25_college_degree df"))
# 
# filter_spline <- filter_rec %>%
#   step_ns(income_per_cap_2019, deg_free = tune("inc_2019 df")) %>%
#   step_ns(income_per_cap_2020, deg_free = tune("inc_2020 df")) %>%
#   step_ns(pct_o25_bachdegree, deg_free = tune("o25_college_degree df"))

trained_rec <- prep(simple_rec, training = train)
trained_data <- bake(trained_rec, new_data = train)

# splined_rec <- prep(simple_spline, training = train)
# splined_data <- bake(splined_rec, new_data = train)


### Old Recipes not currently in use
# simple_norm_rec <- simple_rec %>% 
#   step_normalize(all_numeric_predictors())
# filter_norm_rec <- filter_rec %>% 
#   step_normalize(all_numeric_predictors())
# 
# 
# simple_log_norm_rec <- simple_rec %>% 
#   step_log(all_numeric_predictors(), base = 10) %>% 
#   step_normalize(all_numeric_predictors())
# filter_log_norm_rec <- filter_rec %>% 
#   step_log(all_numeric_predictors(), base = 10) %>% 
#   step_normalize(all_numeric_predictors())

# simple_box_rec <- simple_rec %>% 
#   step_BoxCox(all_numeric_predictors())
# filter_box_rec <- filter_rec %>% 
#   step_BoxCox(all_numeric_predictors())

# makes 0 values
# cannot do box and log together because neither likes nonpositive values

recipe_list <- list(simple_rec = simple_rec,
                    filter_rec = filter_rec,
                    norm_rec = norm_rec)

# recipe_list <- list(simple_norm_rec = simple_norm_rec,
#                     filter_norm_rec = filter_norm_rec, 
#                     simple_log_norm_rec = simple_log_norm_rec, 
#                     filter_log_norm_rec = filter_log_norm_rec)

# make list of workflows
wf_set <- workflow_set(preproc = recipe_list, models = model_list, cross = T)

wf_set <- wf_set %>%
anti_join(tibble(wflow_id = c("filter_rec_linear_reg_2", "spline_filter_linear_reg_2", "filter_rec_rand_forest")),
by = "wflow_id")

```

```{r default recipe workflow pipeline}
set.seed(10)
# TUNING PARAMETERS: try using both regular AND random gridding
wf_metrics <- metric_set(rmse)

# tune each model to the folds and find best parameters (tune_grid) 
# generalize the parameters for now, fine tune them once a general model has been chosen
wf_res <- wf_set %>%
  workflow_map(fn = "tune_grid", resamples = train_split, grid = 5, metrics = wf_metrics)

wf_metric_res <- wf_res %>% collect_metrics()
wf_res %>% autoplot(, select_best = T)
wf_res %>% collect_metrics() %>% arrange(mean)
```


```{r Linear Regression Workflow -- Based on Percentage Pop Recipe}
wf_lin <- workflow() %>% add_model(linear_model) %>% add_recipe(simple_rec)

grid_lin <- grid_regular(parameters(wf_lin), levels = 5)

wf_lin_res <- wf_lin %>% tune_grid(resamples = train_split, grid = grid_lin, metrics = wf_metrics)

wf_lin_metric_res <- wf_lin_res %>% show_best(metric = "rmse", n = 10)

final_wf_param <- wf_lin_res %>% select_best(metric = "rmse")

final_model <- finalize_workflow(wf_lin, final_wf_param)

final_model_fit <- wf_lin %>% fit(data = train)

preds <- final_model_fit %>% predict(new_data = train)


training_pred_chart <- train %>% select(percent_dem) %>% bind_cols(preds)

rmse(truth = percent_dem, estimate = .pred, data = training_pred_chart)

summary(extract_fit_engine(final_model_fit))

# make predictions
preds <- final_model_fit %>% predict(new_data = test)
test_preds <- test %>% select(id) %>% bind_cols(preds) %>% rename(percent_dem = .pred)

# export predictions
write_csv(test_preds, "linearpred.csv")
```


```{r Boosted Tree Workflow}
### Boosted Tree
wf_bt <- workflow() %>% add_model(boost_tree_model) %>% add_recipe(simple_rec)

recipes_param <- extract_parameter_set_dials(boost_tree_model)

grid_bt <- recipes_param %>% grid_regular(levels = 5)

wf_bt_res <- wf_bt %>% tune_grid(resamples = train_split, grid = grid_bt, metrics = wf_metrics)

wf_lin_metric_res <- wf_bt_res %>% show_best(metric = "rmse", n = 10)

final_wf_param <- wf_bt_res %>% select_best(metric = "rmse")

final_model <- finalize_workflow(wf_bt, final_wf_param)

final_model_fit <- final_model %>% fit(data = train)

preds <- final_model_fit %>% predict(new_data = train)

training_pred_chart <- train %>% select(percent_dem) %>% bind_cols(preds)

rmse(truth = percent_dem, estimate = .pred, data = training_pred_chart)

# make predictions
preds <- final_model_fit %>% predict(new_data = test)
test_preds <- test %>% select(id) %>% bind_cols(preds) %>% rename(percent_dem = .pred)

# export predictions
write_csv(test_preds, "boostedpred1.csv")
```


```{r Random Forest Workflow}
### Random Forest -- Eric's
wf_rf <- workflow() %>% add_model(r_forest_model) %>% add_recipe(simple_rec)

recipes_param <- extract_parameter_set_dials(r_forest_model) %>% finalize(trained_data)

grid_rf <- recipes_param %>% grid_regular(levels = 5)

wf_rf_res <- wf_rf %>% tune_grid(resamples = train_split, grid = grid_rf, metrics = wf_metrics)

wf_rf_metric_res <- wf_rf_res %>% show_best(metric = "rmse", n = 10)

final_wf_param <- wf_rf_res %>% select_best(metric = "rmse")

final_model <- finalize_workflow(wf_rf, final_wf_param)

final_model_fit <- final_model %>% fit(data = train)

preds <- final_model_fit %>% predict(new_data = train)

training_pred_chart <- train %>% select(percent_dem) %>% bind_cols(preds)

rmse(truth = percent_dem, estimate = .pred, data = training_pred_chart)

# make predictions
preds <- final_model_fit %>% predict(new_data = test)
test_preds <- test %>% select(id) %>% bind_cols(preds) %>% rename(percent_dem = .pred)

# export predictions
write_csv(test_preds, "rforestpred1.csv")

```

```{r trying out random forest predictions}
### FOR workflow made during zoom -- log_norm_rec not used as based on original simple_rec, so not working currently.
wf_rf <- workflow() %>% add_model(r_forest_model) %>% add_recipe(log_norm_rec)

grid_rf <- grid_regular(parameters(wf_rf), levels = 5)

wf_rf_res <- wf_rf %>% tune_grid(resamples = train_split, grid = grid_rf, metrics = wf_metrics)

wf_rf_metrics <- wf_rf_res %>% show_best(metric = "rmse", n = 5)

final_rf_param <- wf_rf_res %>% select_best(metric = "rmse")

final_rf <- finalize_workflow(wf_rf,final_rf_param)

final_rf_fit <- final_rf %>% fit(data = train)

preds <- final_rf_fit %>% predict(new_data = train)

rf_preds <- train %>% select(percent_dem) %>% bind_cols(preds)

rmse(truth = percent_dem, estimate = .pred, data = rf_preds)

# trying on test data

test_pred_col <- final_rf_fit %>% predict(new_data = test)

test_rf_pred <- test %>% select(id) %>% bind_cols(test_pred_col) %>% rename(percent_dem = .pred)

write_csv(test_rf_pred, "pred3RF.csv")

```
